### 2024-09-27 paper/paper.qmd
update README, paper, references

---
title: "Toronto Island Ferry Tickets: Sales, Redemption Analysis, and Strategic Solutions"
subtitle: ""
author: 
  - Yun Chu
thanks: "code and data supporting this analaysis can be found here: https://github.com/chuyun2024/Analyzing-Toronto-Islands-Ferry-Tickets-Sales-and-Redemption-Count-Data"
date: 26 September 2024
date-format: long
abstract: "This paper presents a comprehensive analysis of the Toronto Islands ferry ticket sales and redemption dataset. The analysis begins with an overview of the dataset, followed by an examination of the cleaned data. In the results section, the relationship between sales and redemption data, the distribution of redemption counts, and the monthly trends in redemption activity are explored. Based on the findings, recommendations to optimize ferry operations, focusing on pricing strategies, staffing allocations, and departure schedules are offered. The paper concludes by discussing the limitations of the analysis and identifying potential areas for future improvement."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
library(tidyverse)
library(knitr)
library(scales)
library(dplyr)
library(kableExtra)
library(ggplot2)
```

# Introduction

Ferries to Toronto Island Park operate year-round, transporting passengers between the Jack Layton Ferry Terminal and Center Island, Hanlan's Point, and Ward's Island, or in the reverse direction from the islands to the terminal. Round-trip tickets are priced at 9.11 Canadian Dollars for adults, with discounted rates available for seniors, youth, juniors, and infants. Currently, the ferries operate on four different schedules, each corresponding to one of the four seasons of the year [@toronto_ferry_website].

These islands are popular destinations for both tourists and residents of Toronto. By analyzing the sales and redemption count data of the ferry service, we can gain insights into visitor trends, helping to optimize resource allocation and revenue management. Identifying peak ferry usage seasons enables a better understanding of when the islands are most popular. This information can be instrumental for governments in resource allocation and operational planning, such as adjusting ferry schedules and staffing levels throughout the year. Additionally, analyzing sales data over time helps uncover revenue trends, providing valuable insights into the financial health of the ferry service and informing pricing strategies.

This paper reveals a strong correlation between sales and redemption data, suggesting a likely linear relationship. It also finds that during the summer months, redemptions are significantly higher compared to winter months. Moreover, 95% of daily redemptions falls below 16,076 @fig-Daily_Redemption_Histogram.

The remainder of this paper is structured as follows. @sec-data1 - @sec-data3 discusses the features of the raw and cleaned data, along with a summary of the cleaned data. @sec-result1 - @sec-result3 investigates the relationship between sales and redemption data using graphs, linear regression, and correlation analysis. It also explores the distribution of redemption counts by frequency and counts. @sec-dis1 - @sec-dis3 further elaborates on the implications and applications of the findings from @sec-result1 - @sec-result3, addresses the limitations of the analysis, and suggests potential improvements.

# Data {#sec-data}

## Raw Data & Measurement{#sec-data1}

The data used in this paper is sourced from the Toronto Open Data portal and was imported using the opendatatoronto library (Gelfand, 2022). The raw dataset contains sales and redemption count data, recorded at 15-minute intervals, collected from both online and POS-kiosk sales. The dataset comprises four variables: ID, timestamp, sales count, and redemption count. The ID variable is a sequential number that records the order of the data entries. Both sales and redemption counts are measured per person, with the redemption count indicating how many individuals successfully redeemed their tickets and boarded the ferries. The sales count represents the total number of tickets sold at kiosks and online. The dataset spans from July 11, 2023, to September 20, 2024, at the time of download [@toronto_ferry_tickets]. Similar datasets were not found online.

The data was downloaded, cleaned, and analyzed using R [@citeR]. Several R packages were utilized throughout the analysis:

- tidyverse [@tidyverse]
- lubridate [@lubridate]
- dplyr [@dplyr]
- knitr [@knitr]
- scales [@scales]
- kableExtra [@kableExtra]

## Cleaned Data{#sec-data2}

The data has been cleaned by removing the hour and minute portions from the Timestamp variable and eliminating the ID variable. Additionally, the sales and redemption count data have been recalculated based on the date rather than the previous 15-minute time intervals.

A sample of the cleaned data (@tbl-cleaned_data) and a line plot (@fig-Sales_and_Redemption_vs_Date) illustrating the sales and redemption counts over time are shown below.

```{r, warning=FALSE}
#| label: tbl-cleaned_data
#| echo: false

# Read the cleaned sales and redemption data from a CSV file
# The file path is relative to the current working directory.
# show_col_types = FALSE suppresses the display of column type information in the output.
data <- read_csv("../data/analysis_data/cleaned_data.csv", show_col_types = FALSE)

# Extract the first few rows of the data for a sample view
# head(data) returns the first 6 rows of the dataset by default
data2 <- head(data)

# Create a formatted table to display the sample data
# kable generates the table, with specified column names and alignment
# 'align = "ccc"' aligns all three columns to the center
# caption adds a descriptive title to the table
kable(data2,
  col.names = c("Date", "Sales", "Redemption"),
  align = "ccc",
  caption = "Sample of Cleaned Sales and Redemption Data"
)
```

```{r, fig.width = 7, fig.height = 4}
#| label: fig-Sales_and_Redemption_vs_Date
#| fig-cap: line plot of sales and redemption data with respect to time
#| echo: false

# Create a line plot using ggplot2
ggplot(data, aes(x = formatted_date)) + # Define the data and aesthetic mappings; formatted_date on the x-axis
  geom_line(aes(y = daily_sales_count, color = "Sales"), linewidth = 1, alpha = 0.7) + # Add a line for daily sales count
  geom_line(aes(y = daily_redemption_count, color = "Redemption"), linewidth = 1, alpha = 0.7) + # Add a line for daily redemption count
  labs(
    title = "Sales and Redemption Over Time", # Add titles and labels to the plot
    x = "Date", # Label for the x-axis
    y = "Count"
  ) + # Label for the y-axis
  scale_color_manual(
    name = "Legend", # Customize the color scale for the legend
    values = c("Sales" = "#2b5b84", "Redemption" = "#98df8a")
  ) + # Assign specific colors to the lines
  theme_minimal() + # Use a minimal theme for cleaner visuals
  theme(plot.title = element_text(hjust = 0.5)) # Center the plot title horizontally
```

The sales and redemption data largely overlap, with significantly higher values in the summer compared to the winter months (@fig-Sales_and_Redemption_vs_Date).

## Summary Statistic{#sec-data3}

To have a better understanding of the data, the minimum, maximum, mean and standard deviation of sales and redemption counts are calculated in @tbl-mean_sales_and_redemption.

```{r}
#| label: tbl-mean_sales_and_redemption
#| echo: false
#| warning: false
#| message: false
#| fig.show: hold

# Calculate summary statistics
summary_statistic <- data |>
  summarise(
    Min_Sales = min(daily_sales_count),
    Max_Sales = max(daily_sales_count),
    Mean_Sales = mean(daily_sales_count),
    Std_Sales = sd(daily_sales_count),
    Min_Redemption = min(daily_redemption_count),
    Max_Redemption = max(daily_redemption_count),
    Mean_Redemption = mean(daily_redemption_count),
    Std_Redemption = sd(daily_redemption_count)
  )

# Create a new data frame for the desired table format
summary_table <- data.frame(
  Min = c(summary_statistic$Min_Sales, summary_statistic$Min_Redemption),
  Max = c(summary_statistic$Max_Sales, summary_statistic$Max_Redemption),
  Mean = c(summary_statistic$Mean_Sales, summary_statistic$Mean_Redemption),
  Standard_Deviation = c(summary_statistic$Std_Sales, summary_statistic$Std_Redemption)
)

# Set row names
rownames(summary_table) <- c("Sales", "Redemption")

# Display the summary table with row and column headings
kable(summary_table,
  format = "latex",
  row.names = TRUE,
  col.names = c("Min", "Max", "Mean", "Standard Deviation"),
  caption = "Summary Statistics for Sales and Redemption Count",
  align = "ccccc"
) |>
  kable_styling(full_width = FALSE, position = "left")

# Note about the table: Setting the format = "latex" resulting that the table does not show under the code chunk, but when rendering, the table shows up fine. If you want to see the table under the code, change format = "html" and it will show up.
```

The summary statistics for sales and redemption counts, presented in @tbl-mean_sales_and_redemption, reveal notable similarities across all metrics.

Given these similarities, it is worthwhile to investigate the relationship between sales and redemption further. This analysis will be explored in @sec-result1.

## Analyais - Investigating the Relationship between the Sales and Redemption Data{#sec-result1}

To investigate the relationship between sales and redemption data, we first create a plot (@fig-Sales_vs_Redemption) that visually describes the relationship between these variables.

```{r,fig.width = 6, fig.height = 3}
#| label: fig-Sales_vs_Redemption
#| fig-cap: Relationship betwwen Sales and Redemption
#| echo: false
#| warning: false
#| message: false

# Create a scatter plot using ggplot2 to visualize the relationship between sales and redemption
ggplot(data, aes(x = daily_sales_count, y = daily_redemption_count)) + # Set the data and aesthetic mappings; daily sales on the x-axis and daily redemption on the y-axis
  geom_point(alpha = 0.5, color = "#008080") + # Add points to the plot to represent each observation
  labs(
    title = "Scatter Plot of Sales vs Redemption of Toronto's Ferries", # Add a title to the plot
    x = "Ticket Sales Count", # Label for the x-axis
    y = "Ticket Redemption Count"
  ) + # Label for the y-axis
  theme(plot.title = element_text(hjust = 0.5))
```

A linear relationship can be seen from the scatter plot. To further investigate this relationship, a linear regression with redemption count as the response variable and sales count as the predictor variable is performed. As shown in @tbl-linear_regression_and_correlation_result, the R-squared statistic is 0.9882, and the correlation coefficient between redemption count and sales count is 0.9941. These values indicate a strong linear correlation between the two variables.

In this paper, we will primarily focus on analyzing the redemption count data, as it directly reflects visitor data. Utilizing redemption counts to analyze revenue is sufficiently accurate due to the high overlap and correlation between the two variables.

```{r}
#| label: tbl-linear_regression_and_correlation_result
#| echo: false
#| warning: false
#| message: false
# Perform linear regression to model the relationship between daily redemption count and daily sales count
model <- lm(daily_redemption_count ~ daily_sales_count, data = data)

# Extract the R-squared value from the summary of the linear model
r_squared <- summary(model)$r.squared

# Calculate the correlation statistic between daily sales count and daily redemption count
corr_stat <- cor(data$daily_sales_count, data$daily_redemption_count)

# Create a summary table to display the R-squared and correlation statistics
summary_table <- data.frame(
  statistic = c("R-Squared", "Correlation"), # Names of the statistics being reported
  value = c(r_squared, corr_stat) # Corresponding values for the statistics
)

# Render the summary table using kable for better formatting, with specified column names and caption
kable(summary_table, col.names = c("Statistics", "Value"), align = "cc", caption = "R-squared and Correlation Statistic")
```

## Analysis - Examining the Frequency of Redemption Count Data{#sec-result2}

@fig-Daily_Redemption_Histogram illustrates that for most days, the redemption count remains below 10,000. The 95 percentile of the redemption count is 16075.4. There are some days where the redemption count peaks at 20,234.

For our resource allocation needs, utilizing monthly data would be sufficient to create an effective staffing schedule. Daily scheduling may not be as feasible or efficient, making a monthly approach more practical for our purposes. The monthly redemption distribution is analysed @sec-result3.

```{r}
#| label: fig-Daily_Redemption_Histogram
#| fig-cap: Histogram of Daily Redemption count data
#| echo: false
#| warning: false
#| message: false
# Calculate the 95th percentile of daily redemption counts
percentile_95 <- quantile(data$daily_redemption_count, 0.95)

# Create a histogram to visualize the distribution of daily redemption counts
ggplot(data, aes(x = daily_redemption_count)) +
  geom_histogram(binwidth = 300, fill = "lightskyblue", color = "steelblue1") + # Add histogram bars with specified binwidth and colors
  labs(
    title = "Histogram of Daily Redemption Counts", # Add a title for the histogram
    x = "Daily Redemption Count", # Label for the x-axis
    y = "Frequency"
  ) + # Label for the y-axis
  theme(plot.title = element_text(hjust = 0.5)) + # Center the plot title
  annotate("text",
    x = percentile_95, y = 30, label = paste("95th Percentile:", round(percentile_95, 2)), # Annotate the plot with the 95th percentile value
    color = "mediumpurple", vjust = -0.5
  ) + # Set the color of the annotation and adjust vertical positioning
  geom_vline(xintercept = percentile_95, color = "mediumpurple", linetype = "dashed", size = 1) # Add a dashed vertical line at the 95th percentile
```

## Analysis - Exploring the Monthly Redemption Data{#sec-result3}

@fig-Redemption_Over_Month illustrates the redemption counts for each month, with dividing lines in each bar indicating daily redemption counts and their cumulative total. The redemption counts are notably low during the winter months, while summer months exhibit significantly higher counts. Notably, in August 2023, the redemption count approached 350,000.

From July to September 2023 and from May to September 2024, the monthly redemption counts exceed 100,000. In contrast, from October 2023 to April 2024, the redemption counts fall below this threshold.

Given these observations, it would be prudent to design two distinct staffing and departure schedules for the ferries: one for peak periods when redemption counts are high and another for off-peak periods when counts are lower. This approach enables more efficient staff and ferries allocation and based on expected demand.

```{r}
#| label: fig-Redemption_Over_Month
#| fig-cap: Redemption Count of Each Month
#| echo: false
#| warning: false
#| message: false
# Create a new column in the data frame to represent the year and month from the formatted date
data <- data %>%
  mutate(year_month = format(formatted_date, "%Y-%m")) # Format the date to "YYYY-MM"

# Create a bar plot to visualize redemption counts over each month
ggplot(data, aes(x = year_month, y = daily_redemption_count)) +
  geom_bar(stat = "identity", fill = "grey93", color = "lightskyblue") + # Create bars with specified fill and border colors
  labs(
    title = "Redemption Over Month", # Add a title for the plot
    x = "Month", # Label for the x-axis
    y = "Redemption Count"
  ) + # Label for the y-axis
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis text for better readability
    plot.title = element_text(hjust = 0.5)
  ) + # Center the plot title
  scale_y_continuous(labels = comma) # Format the y-axis labels to include commas for thousands
```

## Discussion - Impact of Pricing Strategy on Revenue Optimization {#sec-dis1}

The high correlation and linear relationship between sales and redemption data in @sec-result1 suggest that the flat pricing model for ferry tickets has a significant impact. If the fare were adjusted to be lower during the off-peak season, it could potentially enhance revenue during the winter months. While this may result in some revenue loss during the summer, the overall impact might not be as detrimental due to the typically higher volume of sales during that period.

## Discussion - Efficient Staffing Solutions for Peak and Off-Peak Seasons{#sec-dis2}

Based on @fig-Daily_Redemption_Histogram, one measure that Toronto Islands ferries could implement is to integrate staff and departure schedules during peak seasons to prepare for large visitor volumes. For instance, having janitors on-call during busy periods would ensure that adequate cleaning services are available and shortening the time interval between departures when the number of visitors increases significantly.

Additionally, it is essential to develop an on-call schedule for staff that aligns with peak and off-peak seasons. For example, in the winter months, preparing a full staff contingent to manage more than 20,000 visitors per day would be an inefficient use of resources. Instead, a more tailored approach would allow for better resource allocation based on anticipated demand.

## Discussion - Evaluation of Staffing and Departure Schedules in Relation to Visitor Demand{#sec-dis3}

From @fig-Redemption_Over_Month, a noticeable difference in the height of the bars can be observed between May and September compared to the rest of the year. Upon closer examination of the graph, July and August exhibit significantly higher redemption counts compared to other summer months.

Currently, the Toronto Island ferries have adapted four different schedules: Summer (May 16th - September 15th), Fall (September 16th - October 14th), Winter (October 15th - an unspecified date in 2025), and Spring (an unspecified date - an unspecified date in 2025) [@TorontoIslandFerrySchedules]. However, without an analysis of the specific details of these schedules, it remains unclear whether the current allocation of resources is optimal based on the redemption data.

Given the extremely high volume of visitors in July and August, adjusting the staffing and departure schedules during these peak months may be more beneficial. This approach would align more closely with visitor demand and help avoid resource wastage during periods of lower activity.

## Weaknesses and Next Steps

There are a few limitations to address in this analysis.

Firstly, while the examination of the linear relationship between redemption and sales data in @sec-result1 focuses on redemption data for simplicity and as a direct representation of actual visitor volume on the ferries, it is important to note that incorporating sales data into the analysis would yield a more accurate understanding of revenue-related factors. This additional data would provide a clearer and more comprehensive perspective.

Moreover, without a detailed analysis of the current ferry schedules, it is challenging to offer specific recommendations for adjusting these schedules to optimize resource allocation effectively.

\newpage

# References